Cluster Dynamics with HETDEX at z < 0.5 - I: Simulated Performance, Mass Recovery, and Limits to Cosmology

### 2017-01-23 paper1.tex
even more comments

One could use various simulations to attempt to calibrate this observable-mass relation \citeeg{Vanderlinde2010, Sehgal2011}. However, the primary challenge to this method is the incomplete understanding of the baryonic physics which take place in galaxy cluster environments. While there have been (and continue to be) many improvements in the accuracy and power of simulations it is doubtful that in the coming years they will reach the accuracy level required where the observable--mass relation is dominated only by statistics \citep{Weinberg2013}.

The second broad method is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses observations of a relatively small set of clusters and then uses observable mass proxies, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, microwave observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable-mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of cosmic microwave background (CMB) photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a} and can also be effected by the properties of the intracluster medium \citeeg{Pipino2010}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters (where the lensing signal is particularly weak), and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field, such as cluster mergers \citeeg{Evrard2008} and tidal stripping \citeeg{Faltenbacher2006}.

@@

Large-area sky surveys, both on going and planned, are revolutionizing cluster cosmology using a large range of wavelengths. The South Pole Telescope (SPT; \citealt{Carlstrom2011}) and the Atacama Cosmology Telescope (ACT; \citealt{Swetz2011}) are discovering many clusters through the SZE. Optically, the on going Dark Energy Survey (DES; \citealt{DES2005}) and planned Large Synoptic Survey Telescope (LSST; \citealt{LSST2012}) will identify many thousands of clusters to much lower masses than is possible with SZE measurements. However, regardless of the discovery method used, spectroscopic follow-up is needed to further constrain $P(M|z,\vec{x})$. This follow-up becomes increasing important to help constrain the scatter in the mass estimates of other methods, and provides an additional, independent check of the observable-mass relationship used. But as the cluster dataset grows to many tens of thousands of clusters individual follow-up becomes increasingly impractical. Therefore, large spectroscopic surveys are needed to more fully understand the observable-mass relation of clusters.

While a large portion of the $\sim10^6$ interloping lower redshift galaxies will be field (not associated with a bound structure) galaxies, the large area covered by HETDEX is expected to contain as many as 50 Virgo-sized (halo mass $>10^{15}$ \msol) clusters at $z<0.5$. The near-complete spectroscopic coverage allows an unprecedentedly detailed look at a very large number of clusters ranging from group scales to the very massive. In addition to the recovery of accurate dynamical masses, detailed investigations of the of dynamical state of the clusters is possible.

@@

This work plans to address these concerns in the following ways. We create and evaluate a HETDEX like selection ``function'' of galaxies over a similarly large portion of the sky and use well adopted techniques to recover the dynamical properties, such as velocity dispersion and cluster mass. In addition to standard techniques of cluster mass estimation, we investigate probability based and machine learning based approaches of cluster mass prediction. We compare these results to a series of targeted galaxy cluster observations, where each member galaxy is assumed to be observed. Each of these observations use realistic uncertainties from galaxy magnitude and line-flux limits. These strategies will better allow future work to predict the number and types of galaxy clusters which should be observed with VIRUS during both the HETDEX survey portion and through targeted follow up observations.

Throughout this paper, we adopt the following cosmological model: $\Omega_\Lambda = 0.714$, $\Omega_M = 0.286$, $\sigma_8 = 0.82$ and $H_0= 70$ \kms \mpc (taken from the Buzzard catalogs; see below), assume a Chabrier initial mass function (IMF; \citealt{Chabrier2003}), and use AB magnitudes \citep{Oke1974}.

\section{Data and Mock Observations}\label{sec:Data}
In this section, we describe the data products and the techniques used to replicate the HETDEX survey. We use the information from a large mock galaxy catalog enhanced by the emission line properties of galaxies in the SDSS to create a realistic ``sky'' and ``observe'' it with a HETDEX-like observing strategy.

\subsection{The Buzzard Mock Catalogs}\label{sec: Buzzard}
The Buzzard mock galaxy catalogs provide a light cone covering 398.49 \degsq\ between $4^h< RA < 6^h$ and $-61\degree < DEC < -41\degree$ and are derived from a combination of Sub-halo Abundance Matching (ShAM) and ADDSEDs (Adding Density Dependent Spectral Energy Distributions) tied to an in house n-body cosmological simulation. A brief description of the catalog creation is described as follows. The initial conditions are generated with a second-order Lagrangian perturbation theory using {\sc 2LPTic} \citep{Crocce2006}. Dark matter (DM) n-body simulations are run using {\sc LGadget-2} (a version of {\sc Gadget-2}; \citealt{Springel2005}). The Buzzard catalogs adopt the following cosmological parameters: $\Omega_m = 0.286$, $\Omega_\Lambda = 0.714$, $H_0 = 70$ \kms \mpc, $\sigma_8 = 0.82$, and $n_s = 0.96$. The DM halos are identified using the {\sc ROCKSTAR} halo finder \citep{Behroozi2013} which also calculates halo masses.

Galaxy $M_r$ luminosities are added to the velocity peaks using ShAM \citep{Reddick2013}, and ADDSEDs assign luminosities in the other bands. A $M_r$-density-SED relation is created using a SDSS training set, and for each mock galaxy the SED of a randomly selected training set galaxy which has a similar $M_r$ and density is assigned. The assigned SED is also used to determine additional $ugiz$ magnitudes. The result is a mock catalog containing 238 million galaxies with $\sdssr < 29$ mag and $z \leq 8.7$.

@@

\subsection{Conditional {\rm[\ion{O}{II}]} Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS DR12 \citep{Alam2015} catalogs to assign [\ion{O}{II}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 objects classified as galaxies selected over $z = 0.02 - 0.2$ with {\sc zwarning $=0$} and a measured [\ion{O}{II}] line flux signal-to-noise of at least five. Figure~\ref{fig: oii sdss} shows the color-magnitude diagram (CMD) of $M_r$ and $g-r$ for these galaxies, and the following SQL query selects the appropriate galaxies:
\begin{lstlisting}[language=SQL]
SELECT GAL.objid, 
       GAL.u, 
       GAL.err_u, 
       GAL.g, 
       GAl.err_g, 
       GAL.r, 
       GAL.err_r, 
       GAL.i, 
       GAl.err_i, 
       GAL.z, 
       GAL.err_z, 
       g.oii_3726_flux, 
       g.oii_3726_flux_err, 
       s.z 
FROM   galaxy AS GAL 
       JOIN specobj AS s 
         ON s.bestobjid = GAL.objid 
       JOIN galspecline AS g 
         ON g.specobjid = s.specobjid 
WHERE  oii_3726_flux > oii_3726_flux_err * 5 
       AND oii_3726_flux_err > 0 
       AND s.class = 'GALAXY' 
       AND s.zwarning = 0 
\end{lstlisting}

@@

We designed the results of this study to be used in conjunction with HETDEX, a large, blind, spectroscopic survey. HETDEX will measure the redshifts of $8 \times 10^5$ LAE galaxies between $1.9 < z < 3.5$ using a collection of 78 wide-field IFU spectrographs covering the wavelength region $3500 - 5500$~\AAA\ at R $\sim$ 750 \citep{Hill2008}. The primary science goal of these observations is to provide $<1\%$ accuracy measurements of the Hubble expansion parameter and the angular diameter distance at $z\sim2$. This result will provide significant constraints on the evolution of the dark energy equation of state which is both competitive with, and independent of, constraints derived from observations of the \lya forest.

@@

\subsection{Mass estimates}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/Probcomparison.pdf} 
	\caption[Probability based cluster mass predictions.]{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the probability based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The orange, dash-dotted line is the median predicted mass for perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.}
	\label{fig:Probability comparison} 
\end{figure*}

@@

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison.pdf} 
	\caption[ML based cluster mass predictions]{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the ML based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The orange, dash-dotted line is the median predicted mass for perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.}
	\label{fig: ML comparison} 
\end{figure*}

@@

\begin{table*}
\centering
\caption[Mean bias for different bins of predicted cluster mass.]{Mean bias (Eqn.~\ref{eqn: bias}) for different bins of predicted cluster mass. This table shows the bias in the predicted cluster mass for the perfect (top section), Targeted (middle section), and Survey (bottom section) observations in different predicted mass bins. The different mass recovery strategies are given in the leftmost column. It can be used to understand how the predicted cluster mass differs from the true cluster masses. Positive numbers indicate the predicted cluster mass over estimates when compared to the true cluster mass.}
\begin{tabular}{cccccccc} 
		&& \multic{6}{Bins -- Log $M_{pred}$} \\
		\cline{3-8} 
		\multicolumn{2}{c}{Method} & $12.5-13$ & $13-13.5$ & $13.5-14$ & $14-14.5$ & $14.5-15$ & $15-15.5$ \\
		\hline
&& \multic{6}{Perfect Observations} \\
\hline
\multicolumn{2}{c}{Power Law} & $0.23\pm{0.007}$ & $0.16\pm{0.003}$ & $0.11\pm{0.002}$ & $0.07\pm{0.004}$ & $0.02\pm{0.011}$ & $-0.07\pm{0.045}$ \\
\hline 
\rottext{3}{Prob} &$\sigma$ & $0.32\pm{0.005}$ & $0.16\pm{0.003}$ & $0.10\pm{0.002}$ & $0.07\pm{0.004}$ & $0.05\pm{0.012}$ & $-0.18\pm{0.078}$ \\
&$\sigma, z$ & $0.17\pm{0.005}$ & $0.01\pm{0.003}$ & $-0.06\pm{0.002}$ & $-0.11\pm{0.015}$ & $-0.14\pm{0.012}$ & $-0.38\pm{0.159}$ \\
&$\sigma, z, N_{gal}$ & $0.04\pm{0.009}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.002}$ & $-0.05\pm{0.014}$ & $-0.22\pm{0.126}$ & \nd \\
\hline
\rottext{4}{ML} &$\sigma$ & $0.14\pm{0.006}$ & $-0.01\pm{0.003}$ & $-0.07\pm{0.003}$ & $-0.09\pm{0.005}$ & $-0.11\pm{0.014}$ & $-0.17\pm{0.090}$ \\
&$\sigma, z$ & $0.12\pm{0.005}$ & $-0.01\pm{0.003}$ & $-0.06\pm{0.003}$ & $-0.08\pm{0.004}$ & $-0.11\pm{0.012}$ & $-0.24\pm{0.076}$ \\
&$\sigma, N_{gal}$ & $0.04\pm{0.004}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.006}$ & $-0.08\pm{0.041}$ \\
&$\sigma, z, N_{gal}$ & $0.04\pm{0.003}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.001}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.005}$ & $-0.08\pm{0.043}$ \\
\hline

@@

We repeat our analysis using cluster halo and galaxy catalogs from the Millennium simulation obtained via querying the Millennium online database\footnote{\url{http://gavo.mpa-garching.mpg.de/Millennium/}} \citep{Lemson2006}. The Millennium simulation tracks $2160^3$ dark matter particles of $8.6\times 10^8 ~h^{-1}$ \Msol\ inside a comoving 500 $(h^{-1} \mathrm{Mpc})^3$ box from $z=127$ to 0. The following SQL query selects the cluster halos:
\begin{lstlisting}[language=SQL]
SELECT * 
FROM   mfield..fof 
WHERE  snapnum IN (SELECT snapnum 
                   FROM   snapshots..mr 
                   WHERE  z BETWEEN 0.01 AND 0.5) 
       AND m_crit200 > 100 
       AND random BETWEEN 0 AND 10000 
\end{lstlisting}
To select the corresponding galaxies, we use:
\begin{lstlisting}[language=SQL, breaklines=true]
SELECT sh.fofid, 
       sh.subhaloid, 
       g.galaxyid, 
       g.x, 
       g.y, 
       g.z, 
       g.redshift, 
       g.velx, 
       g.vely, 
       g.velz, 
       sdss.g_sdss, 
       sdss.r_sdss, 
       sdss.i_sdss 
FROM   mfield..fofsubhalo AS sh 
       INNER JOIN mpagalaxies..delucia2006a AS g 
              ON sh.subhaloid = g.subhaloid 
       INNER JOIN mpagalaxies..delucia2006a_sdss2mass AS sdss 
              ON g.galaxyid = sdss.galaxyid 
WHERE  sh.fofid IN (SELECT fofid 
       FROM   mfield..fof 
       WHERE  snapnum IN (SELECT snapnum 
                      FROM   snapshots..mr 
                      WHERE  z BETWEEN 0.01 AND 0.5) 
              AND m_crit200 > 1000 
              AND random BETWEEN 0 AND 10000) 
\end{lstlisting}

@@

We conduct our test in two ways. Both use the ML methods (see Section \ref{sec:machine learning method}) to predict the cluster masses of the Millennium clusters, but each test uses a different training set. First, we use the full set of clusters detected in the Buzzard catalogs (14,000 clusters with $M > 10^{11}$ \msol) to train the ML. Second, the Millennium clusters are split into training-testing samples. This provides a test case where we have different cosmological choices between the training and testing samples, and the same cosmological assumptions in both samples. In addition, it provides two different treatments of galaxy evolution in cluster environments. The differing galaxy evolution effects can impact observed galaxy magnitude and color evolution. This provides an important check to understand whether the specific assumptions made in the models affect the results of this analysis.

The top panel of Figure~\ref{fig: mill buzz comparison} shows the ML predicted cluster masses for the $\sim$4000 Millennium clusters as a function of true cluster mass. The orange (Millennium) and green (Buzzard) colors indicate the two different training samples. The median (solid and dashed lines) predicted cluster masses show similar trends regardless of the training data set used. The bottom panel of Figure~\ref{fig: mill buzz comparison} shows the fraction error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The large amount of scatter (the shaded area) in the fractional error for the Buzzard-trained predictions is due to the training set including clusters with masses below the $M = 10^{13}$ \Msol\ threshold for the Millennium clusters. This allows the ML method to predict masses which can be significantly different, whereas the Millennium training set does not include $M < 10^{13}$ \Msol\ clusters, which reduces the scatter of the predicted masses.

Based on these tests, the choice of cosmological parameters does not cause any significant offset in the predicted cluster masses from the trained ML methods. This highlights the versatility of our chosen ML method. The ML method could be further diversified by including cluster measurements from a wide range of cosmological simulations (or observations) which, in affect, marginalizes over all the cosmological assumptions further reducing the dependence.

@@

\subsection{Constraints on Cosmological Parameters}
Galaxy clusters trace the peaks in the universal matter density, often referred to as the power spectrum of matter density fluctuations or the matter power spectrum. This enables them to be sensitive probes of $\Omega_m$, the total mass ($\Omega_b + \Omega_c$) density, and $\sigma_8$, the normalization of the power spectrum. We constrain these parameters by the comparison of the number density of observed clusters to that predicted in cosmological models. Although, in reality, one measures $\sigma_8\Omega_m^q$, where the value of $q$ depends on the masses and redshifts of the halos considered.

>>>>

Previous studies have attempted to use various simulations to calibrate this observable-mass relation \citeeg{Vanderlinde2010, Sehgal2011}. However, the primary challenge to this method is the incomplete understanding of the baryonic physics which take place in galaxy cluster environments. While there have been (and continue to be) many improvements in the accuracy and power of simulations it is doubtful that in the coming years they will reach the accuracy level required, due to limitations of the current models, where the observable--mass relation is dominated only by statistics \citep{Weinberg2013}.

The second broad method is the direct calibration of cluster masses. This recipe has two distinct but often related tracks. The ``direct'' method uses observations of a relatively small set of clusters and then uses observable mass proxies, including X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, microwave observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL; \eg\ \citealt{Rozo2010}) as examples, which provide a ``true'' mass. This directly calibrates the observable-mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are, in fact, estimations, and the methods used to recover these cluster masses are subject to their own limitations. X-ray based cluster masses assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev-Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up-scattering of cosmic microwave background (CMB) photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a} and can also be effected by the properties of the intracluster medium \citeeg{Pipino2010}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters (where the lensing signal is particularly weak), and potentially suffer from line-of-sight effects as WL is sensitive to all mass along the line-of-sight. Virial mass estimators which determine the cluster mass based on the motions of member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of a cluster's potential well, but suffer from systematics due to cluster formation physics which disrupts the velocity field, such as cluster mergers \citeeg{Evrard2008} and tidal stripping \citeeg{Faltenbacher2006}. As all methods possess different strengths, systematics, and assumptions, it is important to compare multiple methods to understand how these affect the inferred cluster mass.

@@

Large-area sky surveys, both on-going and planned, are revolutionizing cluster cosmology using a large range of wavelengths. The South Pole Telescope (SPT; \citealt{Carlstrom2011}) and the Atacama Cosmology Telescope (ACT; \citealt{Swetz2011}) are discovering many clusters through the SZE. Optically, the on going Dark Energy Survey (DES; \citealt{DES2005}) and planned Large Synoptic Survey Telescope (LSST; \citealt{LSST2012}) will identify many thousands of clusters to much lower masses than is possible with SZE measurements. However, regardless of the discovery method used, spectroscopic follow-up is needed to further constrain $P(M|z,\vec{x})$. This follow-up becomes increasingly important to help constrain the scatter in the mass estimates of other methods, and provides an additional, independent check of the observable-mass relationship used. But as the cluster dataset grows to many tens of thousands of clusters individual follow-up becomes increasingly impractical. Therefore, large spectroscopic surveys are needed to more fully understand the observable-mass relation of clusters.

Here, we consider how a blind spectroscopic survey, such as the Hobby Eberly Telescope Dark Energy eXperiment (HETDEX; \citealt{Hill2008}) can be used to calibrate the observable--cluster mass relationship, and to further constrain the scatter in the calibration. We create and evaluate a HETDEX like selection ``function'' of galaxies over a similarly large portion of the sky and use well adopted techniques to recover the dynamical properties, such as velocity dispersion and cluster mass. In addition to standard techniques of cluster mass estimation, we investigate probability based and machine learning based approaches of cluster mass prediction. We compare these results to a series of targeted galaxy cluster observations, where each member galaxy is assumed to be observed. Each of these observations use realistic uncertainties from galaxy magnitude and line-flux limits. These strategies will better allow future work to predict the number and types of galaxy clusters which should be observed with VIRUS during both the HETDEX survey portion and through targeted follow up observations.

Throughout this paper, we adopt the following cosmological model: $\Omega_\Lambda = 0.714$, $\Omega_M = 0.286$, $\sigma_8 = 0.82$ and $H_0= 70$ \kms \mpc (taken from mock galaxy catalogs derived from an N-body simulation using this cosmology; see below), assume a Chabrier initial mass function (IMF; \citealt{Chabrier2003}), and use AB magnitudes \citep{Oke1974}.

@@

\section{Data and Mock Observations}\label{sec:Data}
In this section, we describe the data products and the techniques used to replicate the HETDEX survey. We use the information from a large mock galaxy catalog enhanced by the emission line properties of galaxies detected by the \textit{Sloan Digital Sky Survey} (SDSS; \citealt{Blanton2001a}) to create a realistic ``sky'' and ``observe'' it with a HETDEX-like observing strategy.

\subsection{The Buzzard Mock Catalogs}\label{sec: Buzzard}
The Buzzard mock galaxy catalogs are designed to mimic the sky geometry of DES and provide a light cone covering 398.49 \degsq\ between $4^h< RA < 6^h$ and $-61\degree < DEC < -41\degree$. They are derived from an in-house n-body cosmological simulation. The initial conditions are generated with a second-order Lagrangian perturbation theory using {\sc 2LPTic} \citep{Crocce2006}. Dark matter (DM) n-body simulations are run using {\sc LGadget-2} (a version of {\sc Gadget-2}; \citealt{Springel2005}). The Buzzard catalogs adopt the following cosmological parameters: $\Omega_m = 0.286$, $\Omega_\Lambda = 0.714$, $H_0 = 70$ \kms \mpc, $\sigma_8 = 0.82$, and $n_s = 0.96$. The DM halos are identified using the {\sc ROCKSTAR} halo finder \citep{Behroozi2013} which also calculates halo masses.

Galaxy $M_r$ luminosities are added to the velocity peaks using Sub-halo Abundance Matching (ShAM) \citep{Reddick2013}, and ADDSEDs (Adding Density Dependent Spectral Energy Distributions) assign luminosities in the other bands. A $M_r$-density-SED relation is created using a SDSS training set, and for each mock galaxy the SED of a randomly selected training set galaxy which has a similar $M_r$ and density is assigned. The assigned SED is also used to determine additional $ugiz$ magnitudes. The result is a mock catalog containing 238 million galaxies with $\sdssr < 29$ mag and $z \leq 8.7$.

@@

\subsection{Conditional {\rm[\ion{O}{II}]} Flux Probability Distribution Functions}\label{sec: oii luminosity}
We use the SDSS Data Release 12 \citep{Alam2015} catalogs to assign [\ion{O}{II}] emission line strengths to the galaxies in the Buzzard catalog. We use 503,113 objects classified as galaxies selected over $z = 0.02 - 0.2$ with {\sc zwarning $=0$} and a measured [\ion{O}{II}] line flux signal-to-noise of at least five. Figure~\ref{fig: oii sdss} shows the color-magnitude diagram (CMD) of $M_r$ and $g-r$ for these galaxies, and the following SQL query selects the appropriate galaxies:
\begin{lstlisting}[language=SQL]
SELECT GAL.objid, 
       GAL.u, 
       GAL.err_u, 
       GAL.g, 
       GAl.err_g, 
       GAL.r, 
       GAL.err_r, 
       GAL.i, 
       GAl.err_i, 
       GAL.z, 
       GAL.err_z, 
       g.oii_3726_flux, 
       g.oii_3726_flux_err, 
       s.z 
FROM   galaxy AS GAL 
       JOIN specobj AS s 
         ON s.bestobjid = GAL.objid 
       JOIN galspecline AS g 
         ON g.specobjid = s.specobjid 
WHERE  oii_3726_flux > oii_3726_flux_err * 5 
       AND oii_3726_flux_err > 0 
       AND s.class = 'GALAXY' 
       AND s.zwarning = 0 
\end{lstlisting}

@@

While a large portion of the $\sim10^6$ interloping lower redshift galaxies will be field (not associated with a bound structure) galaxies, the large area covered by HETDEX is expected to contain as many as 50 Virgo-sized (halo mass $>10^{15}$ \msol) clusters at $z<0.5$. The near-complete spectroscopic coverage allows an unprecedentedly detailed look at a very large number of clusters ranging from group scales to the very massive. In addition to the recovery of accurate dynamical masses, detailed investigations of the dynamical state of the clusters is possible.

HETDEX will measure the redshifts of $8 \times 10^5$ LAE galaxies between $1.9 < z < 3.5$ using a collection of 78 wide-field IFU spectrographs covering the wavelength region $3500 - 5500$~\AAA\ at R $\sim$ 750 \citep{Hill2008}. The primary science goal of these observations is to provide $<1\%$ accuracy measurements of the Hubble expansion parameter and the angular diameter distance at $z\sim2$. This result will provide significant constraints on the evolution of the dark energy equation of state which is both competitive with, and independent of, constraints derived from observations of the \lya forest.

@@

\subsection{Mass estimates}
\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/Probcomparison.pdf} 
	\caption[Probability based cluster mass predictions.]{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the probability based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The orange, dash-dotted line is the median predicted mass for Perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.}
	\label{fig:Probability comparison} 
\end{figure*}

@@

\begin{figure*} 
	\includegraphics[width=\textwidth]{figures/MLcomparison.pdf} 
	\caption[ML based cluster mass predictions]{Mass predictions for the power law scaling relation (Equation~\ref{eq:power law}) and the ML based technique with different input features as a function of true cluster mass. The bottom row of panels shows the fractional error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The solid black line shows the 1:1 relation. The orange, dash-dotted line is the median predicted mass for Perfect observations. The solid, purple line is the median predicted mass for the Targeted observing, and the green, dashed line is the median recovered mass for the HETDEX-like observations. The shaded regions represent the 68\% scatter around the median values.}
	\label{fig: ML comparison} 
\end{figure*}

@@

\begin{table*}
\centering
\caption[Mean bias for different bins of predicted cluster mass.]{Mean bias (Eqn.~\ref{eqn: bias}) for different bins of predicted cluster mass. This table shows the bias in the predicted cluster mass for the Perfect (top section), Targeted (middle section), and Survey (bottom section) observations in different predicted mass bins. The different mass recovery strategies are given in the leftmost column. It can be used to understand how the predicted cluster mass differs from the true cluster masses. Positive numbers indicate the predicted cluster mass over estimates when compared to the true cluster mass.}
\begin{tabular}{cccccccc} 
		&& \multic{6}{Bins -- Log $M_{pred}$} \\
		\cline{3-8} 
		\multicolumn{2}{c}{Method} & $12.5-13$ & $13-13.5$ & $13.5-14$ & $14-14.5$ & $14.5-15$ & $15-15.5$ \\
		\hline
&& \multic{6}{Perfect Observations} \\
\hline
\multicolumn{2}{c}{Power Law} & $0.23\pm{0.007}$ & $0.16\pm{0.003}$ & $0.11\pm{0.002}$ & $0.07\pm{0.004}$ & $0.02\pm{0.011}$ & $-0.07\pm{0.045}$ \\
\hline 
\rottext{3}{Prob} &$\sigma$ & $0.32\pm{0.005}$ & $0.16\pm{0.003}$ & $0.10\pm{0.002}$ & $0.07\pm{0.004}$ & $0.05\pm{0.012}$ & $-0.18\pm{0.078}$ \\
&$\sigma, z$ & $0.17\pm{0.005}$ & $0.01\pm{0.003}$ & $-0.06\pm{0.002}$ & $-0.11\pm{0.015}$ & $-0.14\pm{0.012}$ & $-0.38\pm{0.159}$ \\
&$\sigma, z, N_{gal}$ & $0.04\pm{0.009}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.002}$ & $-0.05\pm{0.014}$ & $-0.22\pm{0.126}$ & \nd \\
\hline
\rottext{4}{ML} &$\sigma$ & $0.14\pm{0.006}$ & $-0.01\pm{0.003}$ & $-0.07\pm{0.003}$ & $-0.09\pm{0.005}$ & $-0.11\pm{0.014}$ & $-0.17\pm{0.090}$ \\
&$\sigma, z$ & $0.12\pm{0.005}$ & $-0.01\pm{0.003}$ & $-0.06\pm{0.003}$ & $-0.08\pm{0.004}$ & $-0.11\pm{0.012}$ & $-0.24\pm{0.076}$ \\
&$\sigma, N_{gal}$ & $0.04\pm{0.004}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.006}$ & $-0.08\pm{0.041}$ \\
&$\sigma, z, N_{gal}$ & $0.04\pm{0.003}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.001}$ & $-0.02\pm{0.002}$ & $-0.02\pm{0.005}$ & $-0.08\pm{0.043}$ \\
\hline

@@

We repeat our analysis using cluster halo and \cite{DeLucia2007} galaxy catalogs derived from the Millennium simulation and obtained via querying the Millennium online database\footnote{\url{http://gavo.mpa-garching.mpg.de/Millennium/}} \citep{Lemson2006}. The Millennium simulation tracks $2160^3$ dark matter particles of $8.6\times 10^8 ~h^{-1}$ \Msol\ inside a comoving 500 $(h^{-1} \mathrm{Mpc})^3$ box from $z=127$ to 0. The following SQL query selects the cluster halos:
\begin{lstlisting}[language=SQL]
SELECT * 
FROM   mfield..fof 
WHERE  snapnum IN (SELECT snapnum 
                   FROM   snapshots..mr 
                   WHERE  z BETWEEN 0.01 AND 0.5) 
       AND m_crit200 > 100 
       AND random BETWEEN 0 AND 10000 
\end{lstlisting}
To select the corresponding galaxies, we use:
\begin{lstlisting}[language=SQL, breaklines=true]
SELECT sh.fofid, 
       sh.subhaloid, 
       g.galaxyid, 
       g.x, 
       g.y, 
       g.z, 
       g.redshift, 
       g.velx, 
       g.vely, 
       g.velz, 
       sdss.g_sdss, 
       sdss.r_sdss, 
       sdss.i_sdss 
FROM   mfield..fofsubhalo AS sh 
       INNER JOIN mpagalaxies..delucia2006a AS g 
              ON sh.subhaloid = g.subhaloid 
       INNER JOIN mpagalaxies..delucia2006a_sdss2mass AS sdss 
              ON g.galaxyid = sdss.galaxyid 
WHERE  sh.fofid IN (SELECT fofid 
       FROM   mfield..fof 
       WHERE  snapnum IN (SELECT snapnum 
                      FROM   snapshots..mr 
                      WHERE  z BETWEEN 0.01 AND 0.5) 
              AND m_crit200 > 1000 
              AND random BETWEEN 0 AND 10000) 
\end{lstlisting}

@@

We conduct our test in two ways. Both use the ML methods (see Section \ref{sec:machine learning method}) to predict the cluster masses of the Millennium clusters, but each test uses a different training set. First, we use the full set of clusters detected in the Buzzard catalogs (14,000 clusters with $M > 10^{11}$ \msol) to train the ML. Second, the Millennium clusters are split into training-testing samples. This provides a test case where we have different cosmological choices between the training and testing samples (Buzzard training; Millennium testing), and the same cosmological assumptions in both training and testing. In addition, it provides two different treatments of galaxy evolution in cluster environments. The differing galaxy evolution effects can impact observed galaxy magnitude and color evolution. This provides an important check to understand whether the specific assumptions made in the models affect the results of this analysis.

The top panel of Figure~\ref{fig: mill buzz comparison} shows the ML predicted cluster masses for the $\sim$4000 Millennium clusters as a function of true cluster mass. The orange (Millennium) and green (Buzzard) colors indicate the two different training samples. The median (solid and dashed lines) predicted cluster masses show similar trends regardless of the training data set used. The bottom panel of Figure~\ref{fig: mill buzz comparison} shows the fraction error (Equation~\ref{eq: fractional error}) also as a function of true cluster mass. The large amount of scatter (the shaded area) in the fractional error for the Buzzard-trained predictions is due to the training set including clusters with masses below the $M = 10^{13}$ \Msol\ threshold for the Millennium clusters. This allows the ML method to predict masses which can be significantly different, whereas the Millennium training set does not include $M < 10^{13}$ \Msol\ clusters, which artificially reduces the scatter of the predicted masses.

Based on these tests, the choice of cosmological parameters does not cause any significant offset in the predicted cluster masses from the trained ML methods. While we have only considered physically plausible cosmological models, this highlights the versatility of our chosen ML method. The ML method could be further diversified by including cluster measurements from a wide range of cosmological simulations (or observations) which, in affect, marginalizes over all the cosmological assumptions further reducing the dependence.

@@

\subsection{Constraints on Cosmological Parameters}
Galaxy clusters trace the peaks in the cosmic matter density distribution. This enables them to be sensitive probes of $\Omega_m$, the total mass ($\Omega_b + \Omega_c$) density, and $\sigma_8$, the normalization of the power spectrum. We constrain these parameters by the comparison of the number density of observed clusters to that predicted in cosmological models. Although, in reality, one measures $\sigma_8\Omega_m^q$, where the value of $q$ depends on the masses and redshifts of the halos considered.

To get a sense of how well observations of z < 0.5 galaxy clusters from HETDEX will be able to constrain cosmological parameters we follow the discussion of \defcitealias{Weinberg2013}{W13}\cite{Weinberg2013} (hereafter \citetalias{Weinberg2013}), and begin with a simplifying assumption. We choose to neglect $\Omega_m$ in the following discussion. While sensitive to $\Omega_m$, the number density of clusters does not necessarily provide the strongest constraint; we assume other data sets (\eg, CMB, BAO, supernovae, WL, etc.) will constrain $\Omega_m$ to higher precision and therefore focus solely on $\sigma_8$.
